---
title: "Homework 5"
author: "Mukta Patwari"
date: "2025-10-30"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
library(rvest)

knitr::opts_chunk$set(
  fig.path = "plots/",
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

**Write a function that, for a fixed group size, randomly draws “birthdays” for each person; checks whether there are duplicate birthdays in the group; and returns TRUE or FALSE based on the result.**

```{r}
bday_sim = function(n_room) {
    birthdays = sample(1:365, n_room, replace = TRUE)
    
    repeated_bday = length(unique(birthdays)) < n_room
    
    repeated_bday
  }
```

**Next, run this function 10000 times for each group size between 2 and 50. For each group size, compute the probability that at least two people in the group will share a birthday by averaging across the 10000 simulation runs.**

```{r}
bday_sim_results =
  expand_grid(
    bdays = 2:50,
    iter = 1:10000
  ) %>%
  mutate(
    results = map_lgl(bdays, bday_sim)
  ) %>% 
  group_by(
    bdays
  ) %>% 
  summarize(
    prob_repeat = mean(results)
  )
```

**Make a plot showing the probability as a function of group size, and comment on your results.**

```{r}
bday_sim_results %>% 
  ggplot(aes(x = bdays, y = prob_repeat)) +
  geom_point(color = "red") +
  geom_line() +
  labs(
    x = "Number of people",
    y = "Probability of repeated birthdays"
  )
```

The plot shows that as sample size increases, the probability of repeated birthdays also increases. At around 23 people, the probability of repeated birthdays is 50%, and at around 40 people the probability of repeated birthdays is 87.5%. By the time the sample size is 50, the probability is close to 100%.

## Problem 2

```{r}
library(broom)

sim_t_test = function(n_subj = 30, sigma = 5, mu = 0) {
  
  sim_df =
    tibble(
      x = rnorm(n = n_subj, mean = mu, sd = sigma)
    ) %>% 
    summarize(tidy(t.test(x, mu = 0))) %>%
    select(estimate, p.value, )
}

sim_results_df =
  expand_grid(
    mu = 0:6,
    iter = 1:5000
  ) %>% 
  mutate(
    results = map(mu, ~ sim_t_test(mu = .x))) %>%
  unnest(results)
```

